The evolution of the transcriptome of mammalian placentas

### 2016-11-14 manuscript/manuscript.Rnw
discuss Tlr9/7 response in mice

  \textbf{Results} We identified a set of
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  core genes which are expressed (FPKM $\ge 10$) in all placentas.

  including genes with immune-modulating properties (\textit{ANXA2},
  \textit{ANXA1}, \textit{S100A11}, and \textit{S100A10}) and
  cell-cell interactions (\textit{LAMC1}, \textit{LUM}, and
  \textit{LGALS1}).

@@

Supplemental \cref{tab:shape_genes} shows the top 50 genes which are
associated with the shape of placenta attachment, as well as the shape
in which the difference is most significant. The most significant
genes which were associated with diffuse placenta attachment include
\textit{EREG}, a ligand of EGFR, \textit{UNC93A}, a homolog of
\textit{UNC93B1} which is involved in the balancing between Tlr7 and
Tlr9 immune response in mice\DLA{cite TLR7/TLR9 paper}, and
\textit{WNT5A}, which inhibits the invasion of the trophoblast cell
line HTR8/SVneo \textit{in vitro} \DLA{Cite this paper after reading
  the comments on it (PMID: 26865089)}.

@@

The near exclusivity of deep trophoblast invasion in humans has also
led many to speculate on the origin of various disorders, which are
almost exclusively symptomatic of human pregnancy. For example,
pre-eclampsia is a human disease that occurs in 4\% of pregnancies in
the United States
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. Pre-eclampsia
can be attributed to the lack of adequate invasion of maternal tissue,
and is noted by the onset of hypertension and proteinuria (or in the
absence of proteinuria, thrombocytopenia, impaired liver function,
renal insufficiency, pulmonary edema, or cerebral or visual
disturbances) %
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current,ACOG2013:hypertensionpregnancy}. Some cases of
pre-eclampsia can develop into eclampsia, a more severe condition
resulting in seizures during pregnancy and death. Lack of invasion in
pre-eclampsia may be attributed to the inefficiency of growth
regulators as well as the inability to sufficiently suppress the
maternal immune response. Although there are speculative cases of
eclampsia in great apes
\cite{Baird1981:Eclampsiainlowlandgorilla,Stout.ea1969:Glomerularcapillaryendothelialswelling,Thornton.ea1992:Convulsionsinpregnancyin},
it is exceedingly rare in these species
\cite{Carter.ea2011:Evolutionofinvasiveplacentation}.

>>>>

  \textbf{Results} We identified a set of
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  core genes which are expressed (FPKM $\ge 10$) in all placentas,

  including genes with immune-modulating properties (\textit{ANXA2},
  \textit{ANXA1}, \textit{S100A11}, and \textit{S100A10}), cell-cell
  interactions (\textit{LAMC1}, \textit{LUM}, and \textit{LGALS1}),
  and trophoblast fusion (\textit{ANXA5} and \textit{ANXA1}).

@@

Supplemental \cref{tab:shape_genes} shows the top 50 genes which are
associated with the shape of placenta attachment, as well as the shape
in which the difference is most significant. The most significant
genes which were associated with diffuse placenta attachment include
\textit{EREG}, a ligand of EGFR, \textit{UNC93A}, a homolog of
\textit{UNC93B1} which is involved in the balancing between Tlr7 and
Tlr9 immune response in
mice~\cite{Brinkmann.ea2007:interactionbetweenERmembrane}, and
\textit{WNT5A}, which inhibits the invasion of the trophoblast cell
line HTR8/SVneo \textit{in vitro} \DLA{Cite this paper after reading
  the comments on it (PMID: 26865089)}.

@@

The near exclusivity of deep trophoblast invasion in humans has also
led many to speculate on the origin of various disorders, which are
almost exclusively symptomatic of human pregnancy. For example,
pre-eclampsia is a human disease that occurs in 4\% of pregnancies in
the United States
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. Pre-eclampsia
can be attributed to the lack of adequate invasion of maternal tissue,
and is noted by the onset of hypertension and proteinuria (or in the
absence of proteinuria, thrombocytopenia, impaired liver function,
renal insufficiency, pulmonary edema, or cerebral or visual
disturbances) %
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. Some cases of
pre-eclampsia can develop into eclampsia, a more severe condition
resulting in seizures during pregnancy and death. Lack of invasion in
pre-eclampsia may be attributed to the inefficiency of growth
regulators as well as the inability to sufficiently suppress the
maternal immune response. %
Although there are speculative cases of eclampsia in great apes
\cite{Baird1981:Eclampsiainlowlandgorilla,Stout.ea1969:Glomerularcapillaryendothelialswelling,Thornton.ea1992:Convulsionsinpregnancyin},
it is exceedingly rare in these species
\cite{Carter.ea2011:Evolutionofinvasiveplacentation}.
